Monitoring and Projecting Global Hunger: Are We On Track?

### 2021-03-04 docs/modeling_fies.tex

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[a4paper, margin=1.25in]{geometry}
\usepackage{breakcites}
\usepackage{subcaption}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{textgreek}
\usepackage{authblk}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{lineno}
\usepackage[
  style=numeric,
  citestyle=numeric-comp,
  backend=biber,
  doi=true,
  natbib=true,
  sorting=none
]{biblatex}

@@

@@

\maketitle
\begin{abstract}
This paper presents the first global picture of food security at a subnational level based on the Food Insecurity Experience Scale, the indicator for the Sustainable Development Goal of ``Zero Hunger" that is most indicative of the individual's lived experience of food insecurity and hunger.  Using microdata from 75 countries and filling gaps using machine learning, we find significant heterogeneity in levels of food insecurity around the world.  Examining global temporal trends and accounting for the effects of the COVID-19 pandemic, we find that rates of severe food insecurity are declining, resulting in global decreases in the total number of severely food insecure people.  However, the total number of moderately food insecure people has been increasing and, after recovering from the shock of the COVID-19 pandemic, we predict it will continue to increase through the end of the 2020s.  Overall, we conclude that current trends in development and demographic change will still leave a large share of the world's population still experiencing hunger by 2030.  
\end{abstract}

@@

\begin{table}[H]
  \centering
	\begin{tabular}{lll}
		\toprule
    Name & Source & Scale \\
		\midrule
    Urban Percentage & \citep{Jones2016} & Subnational \\
    Stunting & \citep{Local2020} & Subnational \\
    Wasting & \citep{Local2020} & Subnational \\
    Mean Years of Schooling & \citep{Smits2019, KC2017} & Subnational\\
    GDP Per Capita & \citep{Smits2019, Dellink2017} & Subnational \\
    Gini Coefficient & \citep{Rao2019a} & National\\
    Poverty Headcount Index & \citep{Cuaresma2018} & National \\
    Water Scarcity & \citep{greve2018global} & Subnational \\
    Mean Annual Precipitation &  \cite{abatzoglou2018terraclimate, warszawski2014inter} & Subnational \\
    Topographic Ruggedness &  \cite{USGS1996, Riley1999} & Subnational \\
    Mean Temperature &  \cite{abatzoglou2018terraclimate, warszawski2014inter} & Subnational \\
    Malaria (\textit{P. falciparum}) Mortality Rate &  \cite{Weiss2019} & Subnational \\
		\bottomrule
	\end{tabular}
	\caption{Covariates Included in the Analysis}
	\label{tab:covars}
\end{table}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/TimeSeries.pdf}
  \caption{Number of people over the thresholds for moderate and severe food insecurity, by World Bank income groups as of 2020.  Panel (A) shows the number of food insecure people with region totals stacked, to show global trends and totals over time.  Panel (B) shows regional totals compared against each other over time.}
  \label{fig:timeseries}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/Rates.pdf}
  \caption{Percentage of the population over the threshold for moderate and severe food insecurity, by World Bank income groups as of 2020.}
  \label{fig:rates}
\end{figure}

@@

\begin{landscape}
\begin{figure}[h]
  \centering
  \includegraphics{img/FullMap.pdf}
  \caption{Spatial distribution of moderate and severe food insecurity for the years 2010, 2020, and 2030.}
  \label{fig:map}
\end{figure}
\end{landscape}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{img/VIMP.pdf}
  \caption{Importance of each variable in predicting the percent of a population over thresholds for moderate and severe food insecurity.  This is determined by the increase in the error when the variable is permuted before fitting the model.}
  \label{fig:vimp}
\end{figure}

>>>>

\usepackage{float}
\usepackage[
  colorlinks=true,
  citecolor=blue
]{hyperref}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{authblk}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{makecell}
\usepackage[
  style=authoryear,
  citestyle=authoryear,
  backend=biber,
  doi=true,
  natbib=true,
  sorting=nyt %name, year, title
]{biblatex}

@@

%set up alignment in \makecell
%\renewcommand{\cellalign/theadalign}{cl}

@@

%\maketitle
\begin{abstract}

This paper presents the first global picture of food security at a subnational level based on the Food Insecurity Experience Scale, the indicator for the Sustainable Development Goal of ``Zero Hunger" that is most indicative of the individual's lived experience of food insecurity and hunger.  Using microdata from 75 countries and filling gaps using machine learning, we find significant heterogeneity in levels of food insecurity around the world.  Examining global temporal trends and accounting for the effects of the COVID-19 pandemic, we find that rates of severe food insecurity are declining, resulting in global decreases in the total number of severely food insecure people.  However, the total number of moderately food insecure people has been increasing and, after recovering from the shock of the COVID-19 pandemic, we predict it will continue to increase through the end of the 2020s.  Overall, we conclude that current trends in development and demographic change will still leave a large share of the world's population still experiencing hunger by 2030.  
\end{abstract}

@@

\begin{table}[H]
  \centering
	\begin{tabular}{lll}
		\toprule
    Name & Source & Scale \\
		\midrule
    Urban Percentage & \cite{Jones2016} & Subnational \\ [1cm]
    Stunting & \cite{Local2020} & Subnational \\ [1cm]
    Wasting & \cite{Local2020} & Subnational \\ [1cm]
    Mean Years of Schooling & \makecell[tl]{\cite{Smits2019} \\ \cite{KC2017}} & Subnational\\ [1cm]
    GDP Per Capita & \makecell[tl]{\cite{Smits2019} \\ \cite{Dellink2017}} & Subnational \\ [1cm]
    Gini Coefficient & \cite{Rao2019a} & National\\ [1cm]
    Poverty Headcount Index & \cite{Cuaresma2018} & National \\ [1cm]
    Water Scarcity & \cite{greve2018global} & Subnational \\ [1cm]
    Mean Annual Precipitation & \makecell[tl]{\cite{abatzoglou2018terraclimate} \\ \cite{warszawski2014inter}} & Subnational \\ [1cm]
    Topographic Ruggedness &  \makecell[tl]{\cite{USGS1996} \\ \cite{Riley1999}} & Subnational \\ [1cm]
    Mean Temperature &  \makecell[tl]{\cite{abatzoglou2018terraclimate} \\ \cite{warszawski2014inter}} & Subnational \\ [1cm]
    \makecell[tl]{Malaria (\textit{P. falciparum}) \\ Mortality Rate} & \cite{Weiss2019} & Subnational \\
		\bottomrule
	\end{tabular}
	\caption{Covariates Included in the Analysis}
	\label{tab:covars}
\end{table}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{TimeSeries.pdf}
  \caption{Number of people over the thresholds for moderate and severe food insecurity, by World Bank income groups as of 2020.  Panel (A) shows the number of food insecure people with region totals stacked, to show global trends and totals over time.  Panel (B) shows regional totals compared against each other over time.}
  \label{fig:timeseries}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{Rates.pdf}
  \caption{Percentage of the population over the threshold for moderate and severe food insecurity, by World Bank income groups as of 2020.}
  \label{fig:rates}
\end{figure}

@@

\begin{landscape}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{FullMap.pdf}
  \caption{Spatial distribution of moderate and severe food insecurity for the years 2010, 2020, and 2030.}
  \label{fig:map}
\end{figure}
\end{landscape}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{VIMP.pdf}
  \caption{Importance of each variable in predicting the percent of a population over thresholds for moderate and severe food insecurity.  This is determined by the increase in the error when the variable is permuted before fitting the model.}
  \label{fig:vimp}
\end{figure}
